A weakly structured stem for human origins in Africa

### 2023-03-03 paper.tex
Update paper.tex

Preferring weakly-structured-stem models over archaic-admixture models potentially has 
a range of implications. First, with a weakly structured stem, there is no need to posit 
that an archaic hominin population in Africa stayed reproductively isolated from the 
ancestral human lineage for hundreds of thousands of years before the initiation of 
gene flow. Instead, there would simply have been continuous or recurrent contact between 
two or more groups present in Africa. Additionally, because we would have expected to 
detect admixture from late surviving archaic hominins, the distinctive morphologies of 
certain fairly recent \emph{Homo sapiens} fossils (ADD Harvati et al. 2011 
``The Later Stone Age Calvaria from Iwo Eleru, Nigeria: Morphology and Chronology'', 
Crevecoeur 2012 ``The Upper Paleolithic Human Remains of Nazlet Khater 2 (Egypt) and 
Past Modern Human Diversity'', Crevecoeur et al. 2016 ``Late Stone Age Human Remains 
from Ishango (Democratic Republic of Congo): New Insights on Late Pleistocene Modern 
Human Diversity in Africa''}) may have causes other than admixture \citep{Bergstrom2021-iw.

>>>>

Preferring weakly-structured-stem models over archaic-admixture models potentially has 
a range of implications. First, with a weakly structured stem, there is no need to posit 
that an archaic hominin population in Africa stayed reproductively isolated from the 
ancestral human lineage for hundreds of thousands of years before the initiation of 
gene flow. Instead, there would simply have been continuous or recurrent contact between 
two or more groups present in Africa. Additionally, because we would have expected to 
detect admixture from late surviving archaic hominins, the distinctive morphologies of 
certain fairly recent \emph{Homo sapiens} fossils (ADD Harvati et al. 2011 
``The Later Stone Age Calvaria from Iwo Eleru, Nigeria: Morphology and Chronology'', 
Crevecoeur 2012 ``The Upper Paleolithic Human Remains of Nazlet Khater 2 (Egypt) and 
Past Modern Human Diversity'', Crevecoeur et al. 2016 ``Late Stone Age Human Remains 
from Ishango (Democratic Republic of Congo): New Insights on Late Pleistocene Modern 
Human Diversity in Africa''}) may have causes other than admixture \citep{Bergstrom2021-iw}.
